Generalised gravitational burst generation with Generative Adversarial Networks

### 2020-09-03 Paper_draft/main.tex
Added comments up start of section 4

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{table}[hb]
\centering
\caption{Burst training parameters}
%\footnotesize
\begin{tabular}{@{} l l l l l l }
\br
\hline
 Waveform & Central frequency  & Decay & Central time epoch & Mass range \\
 & (Hz) & (s) & (s) & ($\textrm{M}_{\odot}$) \\
\mr
Sine-Gaussian & 70 - 250 & 0.004 - 0.03 & 0.4 - 0.6 & N/A  \\  
Ringdown & 70 - 250 & 0.004 - 0.03 & 0.4 - 0.6 & N/A \\
White-noise burst & 70 - 250 & 0.004 - 0.03 & 0.4 - 0.6 & N/A  \\
Gaussian pulse & N/A & 0.004 - 0.03 & 0.4 - 0.6 & N/A  \\
BBH & N/A & N/A & N/A & 5 - 70  \\
 \br
\end{tabular}\\
\label{Tab:training_parms}
\end{table}
\normalsize

GW burst signals remain an unmodelled phenomenon, as such, current
detection algorithms focus on waveforms signals appearing within multiple detectors We propose a
signal generation scheme using \acp{GAN} trained on burst-like
waveforms. BurstGAN is a conditional GAN that is trained on five signal morphology's spanning a range of prior
parameters. The parameter ranges can be seen in \cref{Tab:training_parms}. The families are:

\begin{itemize}

\item Sine-Gaussian $h_{sg}(t) = A \exp\left[ - (t-t_{0})^2 / \tau^2 \right] \sin (2 \pi f_0 (t-t_0))$, a sinusoidal wave with a Gaussian envelop characterised by a central frequency, $f_0$;.
\item Ring-down, $h(t) = A \exp \left[-{(t-t_0)} / {\tau} \right] \sin(2 \pi f_0 (t-t_0))$, with frequency $f_0$ and duration $\tau$
\item White-noise bursts, with bounded frequency bandwidth $\Delta f$ and duration $\tau$.
\item Gaussian pulse $h(t) = \exp(-t^2 / \tau^2)$ with duration. $\tau$.
\item Binary-black hole inspirals which are simulated using IMRPhenomD waveform \cite{Khan_2016} routine from
LALSuite \cite{lalsuite} which models the
inspiral, merger and ringdown of a \ac{BBH} waveform. The component masses lie
in the range of [5,70] $\textrm{M}_{\odot}$ with zero spins
and we fix m$_1$ $>$ m$_2$. The mass distribution is approximated by a power
law with index of 1.6 \cite{Abbott_2019}. The signals are generated using random right ascensions and
declinations uniform over the sky and the inclinations are drawn from the
cosine of a uniform distribution in the range [-1,1]. 
\end{itemize}
The peaks of the waveforms are set to be within [0.4, 0.6]s of the 1s time interval and all training waveforms are sampled at 1024 Hz.

\subsection{Architecture details}

% Introduce some architecture features
%
Extensions to the original \ac{GAN} method such as the
\ac{DCGAN}~\cite{Radford2015} have been widely praised for
enabling a stable \ac{GAN} architecture. Most GANs now replace fully connected layers, where each neuron is connect to all the neurons in the previous and next layer, with convolutional layers. \ac{CNN} \michael{capitalisation, not sure how it works with the accronyms package} are designed to work with grid-like structures with close local dependencies like image and text based data, however, there are examples of audio synthesis work \cite{DBLP:journals/corr/abs-1809-11096}.

% specifically define our architecture
%
We adopt the suggestions of \cite{Radford2015,DBLP:journals/corr/abs-1809-11096}, lengthening
one-dimensional convolution kernels on both the generator and discriminator.
The Generator model is fully convolutional, upsampled using strided transposed
convolutions with batch normalisation in the first layer and ReLU activations
throughout with the exception of Tanh for the output layer. Each transposed
convolutional layer uses a kernel size of $18\times 1$ and stride of 2. The
discriminator network mirrors that of the generator without batch
normalization, using LeakyReLU activations, SpatialDropout, and a 2-stride
convolution for downsampling. The discriminator output is a single node activated by a Sigmoid that can be interpreted as
a probability of the the signal being real. This model is trained
with binary cross entropy. The full architecture description can be seen in \cref{Tab:hyperparameters}. ~\jordan{still too technical im not sure if we need to go into more detail on each thing mentioned.} \michael{This para doesn't seem to technical to me}
%~\chris{OK, very technically tight but too
%computer sciencey. Try to make it human readbale for a physicist and refer to
%the table of parameters in the appendix.}

% describe the hyperparameter tunings
%
Neural networks and subsequently \acp{GAN} have multiple parameters a developer
can tune when designing the model and these are referred to as hyperparameters.
The final network design used in this work comes from the use of trial and
error and the initial designs influenced by the available literature. We found that the GAN performed better with both networks having the same number of layers and neurons which should encourage even competition between the generator and discriminator.  After
tuning the multiple hyperparamters (\cref{Tab:hyperparameters}), the GAN
was trained on $10^5$ signals which contained an even mixture of sine-gaussian, ringdown, white noise bursts, gaussian blips and BBHs for 1000 epochs and takes $\Or$(2) days to train. A larger batch size saw a small increase in performance at the cost of training speed and we decided that 128 was a fair compromise.

@@

~\chris{In general I find this section to read like un-connected statements in
the sense that they are un-connected to each other and un-connected to the main
GW problem. Try to make it more like a recipe for doing this analysis.}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{table}[hb]
\centering
\caption{Burst training parameters~\chris{Add a bit more caption description
here. Plus address the issues of the actual distributions used (uniform) and
how you include the BBH parameters. It's also not obvious that all the classes
share the same 3 parameters.}}
%\footnotesize
\begin{tabular}{@{} l l l l l l }
\br
\hline
 Waveform & Central frequency  & Decay & Central time epoch & Mass range \\
 & (Hz) & (s) & (s) & ($\textrm{M}_{\odot}$) \\
\mr
Sine-Gaussian & 70 - 250 & 0.004 - 0.03 & 0.4 - 0.6 & N/A  \\  
Ringdown & 70 - 250 & 0.004 - 0.03 & 0.4 - 0.6 & N/A \\
White-noise burst & 70 - 250 & 0.004 - 0.03 & 0.4 - 0.6 & N/A  \\
Gaussian pulse & N/A & 0.004 - 0.03 & 0.4 - 0.6 & N/A  \\
BBH & N/A & N/A & N/A & 5 - 70  \\
 \br
\end{tabular}\\
\label{Tab:training_parms}
\end{table}
\normalsize

% introduce the signal models
%
\ac{GW} burst signals remain an unmodelled phenomenon~\chris{as in the
introduction I think you need to make the distinction between unmodelled and
modelled burst analyses}, as such, current detection algorithms exploit the
fact that astrophysical signals will appear within data from multiple detectors
in coincidence and share the same waveform morphology.~\chris{I have added to
this first sentence but it still doesn't naturally lead intpo the next one} We
propose a signal generation scheme using a \ac{CGAN} trained on burst-like
waveforms. We call this \texttt{BurstGAN}~\chris{give reference to code repo
here} and is a \ac{CGAN} trained on five signal classes each spanning a range
of prior signal parameters. The signal classes are:

% list the 5 waveform classes
%
\begin{itemize}
%
\item {\bf Sine-Gaussian}: $h_{\text{sg}}(t) = A \exp\left[ - (t-t_{0})^2 /
\tau^2 \right] \sin (2 \pi f_0 (t-t_0))$, a sinusoidal wave with a Gaussian
envelope characterised by a central frequency, $f_0$.~\chris{what about
amplitude, time of arrival and initial phase? For all the other classes too.}
%
\item {\bf Ring-down}: $h_{\text{rd}}(t) = A \exp \left[-{(t-t_0)} / {\tau}
\right] \sin(2 \pi f_0 (t-t_0))$, with frequency $f_0$ and duration $\tau$
%
\item {\bf White-noise bursts}: $h_{\text{wn}}(t_j) = Ag_j\exp\left[ -
(t-t_{0})^2 / \tau^2 \right]$ where $g_j$ are drawn from a zero mean unit
variance Gaussian distribution and ahave a bounded frequency bandwidth $\Delta
f$ and duration $\tau$~\chris{no, they have a Gaussian envelope with a duration
parameter $tau$, we don't explicitly define the bandwidth}.
%
\item {\bf Gaussian pulse}: $h_{\text{gp}}(t) = \exp(-t^2 / \tau^2)$ with
duration parameter $\tau$.
%
\item {\bf Binary-black hole}: Simulated using the IMRPhenomD
waveform~\cite{Khan_2016} routine from LALSuite~\cite{lalsuite} which models
the inspiral, merger and ringdown of a \ac{BBH} waveform. The component masses
lie in the range of [5,70] $\textrm{M}_{\odot}$ with zero spins and we fix
$m_1>m_2$. The mass distribution is approximated by a power law with
index of 1.6~\cite{Abbott_2019}. The signals are generated using random right
ascensions and declinations uniform over the sky and the inclinations are drawn
from the cosine of a uniform distribution in the range [-1,1].~\chris{OK, we
have an issue here. First the prior parameters and ranges are being defined
here rather than in the table with the other classes. Second, you are
randomising the sky in order to get random antenna patterns which only act to
change the relative amplitudes in multiple detectors - whcih we don't use any
more. So I would say that you shoudl only use the plus polarisation since it is
the same as the cross but pi/2 out of phase.}  
%
\end{itemize}
%
The location of the peak amplitude of the waveforms (corresponding to the
mid-points of all but the ring-down and \ac{BBH} classes) are randomly drawn to
be within [0.4, 0.6] sec from the start of the 1 sec time interval and all
training waveforms are sampled at 1024 Hz.  The parameter prior ranges are
defined in~\cref{Tab:training_parms}.~\chris{Yes, those are the ranges but what
are the distributions?}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Architecture details}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Introduce some architecture features
%
Extensions to the original \ac{GAN} method such as the
\ac{DCGAN}~\cite{Radford2015} have been widely praised for enabling a stable
\ac{GAN} architecture. Most GANs now replace fully connected layers, where each
neuron is connect to all the neurons in the previous and next layer, with
convolutional layers. \acp{CNN}~\michael{capitalisation, not sure how it works
with the accronyms package}~\chris{I will fix this} are designed to work with
grid-like structures with close local dependencies like image and text based
data, however, there are examples of audio synthesis
work~\cite{DBLP:journals/corr/abs-1809-11096}.~\chris{this text sounds very
similar to the CNN text from earlier. Do you need this here? Can you simply say
that we are using a DCGAN architecture or say that our networks use
convolutional layers and then move on?}

% specifically define our architecture
%
We adopt the suggestions
of~\cite{Radford2015,DBLP:journals/corr/abs-1809-11096}, lengthening
one-dimensional convolution kernels on both the generator and
discriminator~\chris{what does this statment actually mean and why are you
opening the paragrpah with it?}.  The Generator model is fully convolutional,
upsampled using strided transposed~\chris{you mentioned strides before but not
transposing. Can you explain that aspect seomwhere earlier on.} convolutions
with batch normalisation~\chris{same here, can you mention it earlier} in the
first layer and ReLU~\chris{you can easily mention ReLU and other activations
earlier} activations throughout with the exception of the Tanh activation for
the output layer~\chris{why do you need this for the output? State that it is
to guarantee the output function is bound between [-1,1]}. Each transposed
convolutional layer uses a kernel size of $18\times 1$~\chris{by saying 18x1
you are saying that you are using a 2D kernel with width 1 but why not just say
you use a 1D kernel of size 18?} and stride of 2. The discriminator network
mirrors that of the generator without batch normalization~\chris{why?}, using
LeakyReLU activations, SpatialDropout, and a 2-stride convolution for
downsampling~\chris{any reason for these choices?}. The discriminator output is
a single node with Sigmoid activation that can be interpreted as a probability
of the the signal being real. This model is trained with binary cross
entropy~\chris{how exactly if the output is 1 number rather than 2 (binary)?}.
The full architecture description can be seen in~\cref{Tab:hyperparameters}.
~\jordan{still too technical im not sure if we need to go into more detail on
each thing mentioned.} \michael{This para doesn't seem to technical to
me}~\chris{same here, not too technical but lacking detail and motivation for
choices.}
%~\chris{OK, very technically tight but too computer sciencey. Try to make it
%human readbale for a physicist and refer to the table of parameters in the
%appendix.}

% describe the hyperparameter tunings
%
Neural networks and subsequently \acp{GAN} have multiple parameters a developer
can tune when designing the model and these are referred to as hyperparameters.
The final network design used in this work comes from the use of trial and
error and the initial designs influenced by the available literature. We found
that the \ac{GAN} performed better with both networks having the same number of
layers and neurons which should~\chris{should? well does it?} encourage even
competition between the generator and discriminator.  After tuning the multiple
hyperparameters (see \cref{Tab:hyperparameters}), the \ac{GAN} was trained on
$10^5$ signals which contained an even mixture~\chris{be a bit more specific
about what an even mixture is - technically they are drawn from a categorical
distribution with equal propabilities for each class} of
sine-Gaussian~\chris{Gauss was a person so we have to capitalise his stuff},
ring-down, white noise bursts, Gaussian pulse and \acp{BBH} for 1000 epochs and
takes $\Or$(2) days to train~\chris{on what? be specific}. A larger batch size
saw a small increase in performance~\chris{what is performance in this case} at
the cost of training speed and we decided that 128~\chris{128 what? Also, why
pick on batch size as the thing to comment on here? What about learning rate,
kernel size, etc...} was a fair compromise.

@@

~\chris{Please add a loss plot and use a paragrpah to describe what is going
on.}
